---
title: "Untitled"
output: html_document
---

pairwise differential GSVA score testing (using limma) of secRecon geneset activity for transcriptome and proteome plasma vs CHO data

refer to 1_correlation_hclust.Rmd for generation of GSVA scores
```{r}
library(limma)
library(tidyr)
library(dplyr)
library(EnhancedVolcano)
```

```{r}
# load in GSVA and data image from previous notebook
# load("../data/1_correlation_hclust.RData")
```


how to interpret GSVA scores and compare
https://support.bioconductor.org/p/119709/

vignette tutorial to apply LIMMA to find differential pathway activity
https://www.bioconductor.org/packages/release/bioc/vignettes/GSVA/inst/doc/GSVA.html#62_Differential_expression_at_pathway_level


```{r}
# define sample groups/labels
rna.groups <- list()
rna.groups$cho <- c(CHO_DG44.t,CHO_K1.t)
rna.groups$mm <- c(MPC.t,P3X.t)
rna.groups$hu <- c(JK.t,Karpas.t)
rna.groups$plasma = c(rna.groups$mm,rna.groups$hu)

prot.groups <- list()
prot.groups$cho <- c(CHO_DG44.p,CHO_K1.p)
prot.groups$mm <- c(MPC.p,P3X.p)
```

# generate limma models
```{r}
# generate model (comparison: mus plasma vs CHO)
mod.meta.factor <- as.factor(c(rep('mm',6),rep('cho',6)))
mod <- model.matrix(~ factor(mod.meta.factor))
mod
colnames(mod) <- c("int", "MMvsCHO")
mod
```

```{r}
# generate models for intraspecies analyses
## rna:
### human: JK vs Karpas
### mu: MPC vs P3X63
### CHO: CHO-DG44 vs CHO-K1

# rna.groups$cho <- c(CHO_DG44.t,CHO_K1.t)
# rna.groups$mm <- c(MPC.t,P3X.t)
# rna.groups$hu <- c(JK.t,Karpas.t)

## prot:
### mu: MPC vs P3X63
### CHO: CHO-DG44 vs CHO-K1

cl.comparisons = list('cho'='CHO-DG44.vs.CHO-K1',
                      'mm'='MPC.vs.P3X63',
                      'hu'='JK.vs.Karpas')

mod.meta.factor.cl <- as.factor(c(rep('CL2',3),rep('CL1',3)))
mod.cl <- model.matrix(~ factor(mod.meta.factor.cl))
mod.cl
colnames(mod.cl) <- c("int", "CL2vsCL1")
mod.cl
```


# fit limma models
```{r}
gsva.limma.fit <- list()
gsva.limma.fit$rna <- eBayes(lmFit(cho.plasma.gsva[,c(rna.groups$mm,rna.groups$cho)], mod))
gsva.limma.fit$pro <- eBayes(lmFit(cho.plasma.gsva.pro, mod))

gsva.limma.res <- lapply(gsva.limma.fit,function(x) decideTests(x, p.value=0.05))

lapply(gsva.limma.res,summary)

```

```{r}
# fit models for intraspecies comparisons
gsva.limma.fit.cl <- list()

for (species in c('cho','mm','hu')) {
  gsva.limma.fit.cl[[paste('rna',cl.comparisons[[species]],sep='_')]] <- eBayes(lmFit(cho.plasma.gsva[,rna.groups[[species]]], mod.cl))
}

for (species in c('cho','mm')) {
  gsva.limma.fit.cl[[paste('pro',cl.comparisons[[species]],sep='_')]] <- eBayes(lmFit(cho.plasma.gsva.pro[,prot.groups[[species]]], mod.cl))
}

gsva.limma.res.cl <- lapply(gsva.limma.fit.cl,function(x) decideTests(x, p.value=0.05))

lapply(gsva.limma.res.cl,summary)
```

# limma results
refer to GSVA vignette on interpretation of logFC as difference in activity, where the magnitude difference and stringency should not be directly interpreted like expression fold-change
https://www.bioconductor.org/packages/release/bioc/vignettes/GSVA/inst/doc/GSVA.html#62_Differential_expression_at_pathway_level

```{r}
# extract limma results (toptable) for transcriptomics and proteomics GSVA; separate into up and downregulated using adjusted p-value significance (0.05)
gsva.limma.tt <- lapply(gsva.limma.fit,function(x) {topTable(x,coef=2,number=length(sec.db))})

gsva.limma.tt$rna.up <- subset(gsva.limma.tt$rna,logFC > 0 & adj.P.Val<0.05)
gsva.limma.tt$rna.down <- subset(gsva.limma.tt$rna,logFC < 0 & adj.P.Val<0.05)
gsva.limma.tt$pro.up <- subset(gsva.limma.tt$pro,logFC > 0 & adj.P.Val<0.05)
gsva.limma.tt$pro.down <- subset(gsva.limma.tt$pro,logFC < 0 & adj.P.Val<0.05)
```

```{r}
# overall significant differences in secRecon subprocess activity
gsva.limma.sig <- list()
gsva.limma.sig$rna <- subset(gsva.limma.tt$rna, adj.P.Val<0.05)
gsva.limma.sig$pro <- subset(gsva.limma.tt$pro, adj.P.Val<0.05)
```


```{r}
# extract significant results for intraspecies cell lines
gsva.limma.tt.cl <- lapply(gsva.limma.fit.cl, function(x) {topTable(x,coef=2,number=length(sec.db))})
  
gsva.limma.sig.cl <- lapply(gsva.limma.tt.cl, function(x) {
  subset(x,adj.P.Val<0.05)})

gsva.limma.sig.cl.up <- lapply(gsva.limma.sig.cl, function(x) {
  subset(x,logFC > 0)})

gsva.limma.sig.cl.down <- lapply(gsva.limma.sig.cl, function(x) {
  subset(x,logFC < 0)})

gsva.limma.sig.cl
```


identify differential process at both RNA and protein level
```{r}
intersect(rownames(gsva.limma.tt$rna.up),rownames(gsva.limma.tt$pro.up))
intersect(rownames(gsva.limma.tt$rna.down),rownames(gsva.limma.tt$pro.down))
```

intraspecies -- differential terms for rna and protein
```{r}
intersect(rownames(gsva.limma.sig.cl$`rna_CHO-DG44.vs.CHO-K1`),
          rownames(gsva.limma.sig.cl$`pro_CHO-DG44.vs.CHO-K1`)) # 28 terms (3 are from Feizi db)

# intersect(rownames(gsva.limma.sig.cl$rna_MPC.vs.P3X63),
#           rownames(gsva.limma.sig.cl$`pro_MPC.vs.P3X63`))
# "Lysosomal degradation/Glyco-conjugate degradation"
```


# visualize differential GSVA pathways
generate volcano plots


```{r}
# interspecies CHO vs plasma
# define secRecon processes differential at rna and protein level
sr.process.ev <- c('ERpQC', 'Vesicle budding', 'PERK pathway', 'Co-translational translocation', 'N-glycan','Hydroxylysine')
```

```{r}
# at intraspecies/cell line level
sr.process.ev.cl <- list(
  'cho' = grep('f_',c(
    intersect(rownames(gsva.limma.sig.cl.up$`rna_CHO-DG44.vs.CHO-K1`),rownames(gsva.limma.sig.cl.up$`pro_CHO-DG44.vs.CHO-K1`)),
    intersect(rownames(gsva.limma.sig.cl.down$`rna_CHO-DG44.vs.CHO-K1`),rownames(gsva.limma.sig.cl.down$`pro_CHO-DG44.vs.CHO-K1`)))
    ,invert = T,value = T),
  'mu' = intersect(rownames(gsva.limma.sig.cl$rna_MPC.vs.P3X63),
                   rownames(gsva.limma.sig.cl$`pro_MPC.vs.P3X63`)))
```



```{r}
# assign unique colorpalette to metadata
## sec.meta object defined in previous notebook

# define a named vector of process of interest : color
sec.meta.colors.pg <- c('blue3','orange','green','darkkhaki','skyblue')

names(sec.meta.colors.pg) = grep('f_',unique(sec.meta$sec.parents),invert = T,value = T)
sec.meta.colors.pg

```

RNA GSVA volcano plot
```{r}
# create color legend
ev.df <- gsva.limma.tt$rna

# remove feizi rows
ev.df <- ev.df[!grepl('f_',rownames(ev.df)),]

ev.df.col <- c()
for (i in rownames(ev.df)) {
  if (ev.df[i,'adj.P.Val'] > 0.05) {
    ev.df.col <- append(ev.df.col,'grey')
  } else {
    i.p <- sec.meta[i,'sec.parents']
    ev.df.col <- append(ev.df.col,sec.meta.colors.pg[[i.p]])
}}
names(ev.df.col) <- rownames(ev.df)
```

```{r}
# create color legend for intraspecies cell line plots
ev.df.cl <- list('cho'=gsva.limma.tt.cl$`rna_CHO-DG44.vs.CHO-K1`,
                 'mm'=gsva.limma.tt.cl$rna_MPC.vs.P3X63)


ev.df.cl.viz <- lapply(ev.df.cl,function(x) {
  x.sr <- x[!grepl('f_',rownames(x)),] # remove feizi rows
  print(dim(x.sr))
  # print(head(rownames(x.sr)))
  print(head(x.sr))
  x.col <- c()
  # return(x.sr)
  for (i in rownames(x.sr)) {
    print(i)
    print(x.sr[i,'adj.P.Val'])
    if (x.sr[i,'adj.P.Val'] > 0.05) {x.col <- append(x.col,'grey')} else {i.p <- sec.meta[i,'sec.parents']
      x.col <- append(x.col,sec.meta.colors.pg[[i.p]])}}
  names(x.col) <- rownames(x.sr)
  return(list('data'=x.sr,'color'=x.col))
  })

# ev.df.cl.viz$cho$color
```


volcano plot for RNA- PCD (murine plasma) vs CHO
```{r fig.height=4, fig.width=5}
p <- EnhancedVolcano(ev.df,
    lab = rownames(ev.df),
    colCustom = ev.df.col,
    selectLab = sr.process.ev,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(ev.df$logFC),max(ev.df$logFC)),
    ylim=c(0,-log10(min(ev.df$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle('GSVA - PCD vs CHO (RNA)') + theme(legend.position="none",plot.title = element_text(hjust = 0.5)) #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'right')

## uncomment above lines to create a manual legend for fianl figures

ggsave("gsva_rna.svg", plot = p, width = 10, height = 8)
```

volcano plot for RNA- CHO intraspecies
```{r fig.height=4, fig.width=5}
df.viz <- ev.df.cl.viz$cho$data
df.colors <- ev.df.cl.viz$cho$color
p <- EnhancedVolcano(df.viz,
    lab = rownames(df.viz),
    colCustom = df.colors,
    selectLab = sr.process.ev.cl$cho,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(df.viz$logFC),max(df.viz$logFC)),
    ylim=c(0,-log10(min(df.viz$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle(paste('GSVA -',cl.comparisons$cho,'(RNA)')) + theme(legend.position="none",plot.title = element_text(hjust = 0.5)) #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'right')

## uncomment above lines to create a manual legend for fianl figures

print(p)
# ggsave("cho_gsva_rna.svg", plot = p, width = 10, height = 8)
```

volcano plot for RNA- murine intraspecies
```{r fig.height=4, fig.width=5}
df.viz <- ev.df.cl.viz$mm$data
df.colors <- ev.df.cl.viz$mm$color
p <- EnhancedVolcano(df.viz,
    lab = rownames(df.viz),
    colCustom = df.colors,
    selectLab = sr.process.ev.cl$mu,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(df.viz$logFC),max(df.viz$logFC)),
    ylim=c(0,-log10(min(df.viz$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle(paste('GSVA -',cl.comparisons$mm,'(RNA)')) + theme(legend.position="none",plot.title = element_text(hjust = 0.5)) #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'right')

## uncomment above lines to create a manual legend for fianl figures

print(p)
# ggsave("mm_gsva_rna.svg", plot = p, width = 10, height = 8)
```

PROTEIN GSVA volcano plot
```{r}
# create color legend for CHO vs murine plasma
ev.df <- gsva.limma.tt$pro

# remove feizi rows
ev.df <- ev.df[!grepl('f_',rownames(ev.df)),]

ev.df.col <- c()
for (i in rownames(ev.df)) {
  if (ev.df[i,'adj.P.Val'] > 0.05) {
    ev.df.col <- append(ev.df.col,'grey')
  } else {
    i.p <- sec.meta[i,'sec.parents']
    ev.df.col <- append(ev.df.col,sec.meta.colors.pg[[i.p]])
}}
names(ev.df.col) <- rownames(ev.df)
```

```{r}
# create color legend for intraspecies cell line viz
ev.df.cl <- list('cho'=gsva.limma.tt.cl$`pro_CHO-DG44.vs.CHO-K1`,
                 'mm'=gsva.limma.tt.cl$pro_MPC.vs.P3X63)


ev.df.cl.viz <- lapply(ev.df.cl,function(x) {
  x.sr <- x[!grepl('f_',rownames(x)),] # remove feizi rows
  x.col <- c()
  for (i in rownames(x.sr)) {
    if (x.sr[i,'adj.P.Val'] > 0.05) {x.col <- append(x.col,'grey')} else {i.p <- sec.meta[i,'sec.parents']
      x.col <- append(x.col,sec.meta.colors.pg[[i.p]])}}
  names(x.col) <- rownames(x.sr)
  return(list('data'=x.sr,'color'=x.col))
  })

# ev.df.cl.viz$cho$color
```

volcano for CHO vs murine plasma
```{r fig.height=4, fig.width=5}
EnhancedVolcano(ev.df,
    lab = rownames(ev.df),
    colCustom = ev.df.col,
    selectLab = sr.process.ev,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(ev.df$logFC),max(ev.df$logFC)),
    ylim=c(0,-log10(min(ev.df$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle('GSVA - PCD vs CHO (Protein)') + theme(legend.position="none",plot.title = element_text(hjust = 0.5))
     #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'left')

###### UNCOMMENT above lines to create a manual legend for fianl figures
 
#   geom_rect(aes(xmin = -0.21, xmax = -0.1, ymin = 5, ymax = 5.5), fill = "white", color = "black") +
#   theme(plot.title = element_text(hjust = 0.5),  # Center the title
#       legend.position = "none") +  # Remove default legend
#   annotate("text", x = -0.21, y = 5.8, label = "System", fontface = "bold", hjust = 0,size = 6)
# annotate("text", x = -0.21, y = 5.7, label = "â— Proteostasis", color = "blue3", hjust = 0) 

ggsave("gsva_prot.svg", plot = p, width = 10, height = 8)
```

volcano plot for RNA- CHO intraspecies
```{r fig.height=4, fig.width=5}
df.viz <- ev.df.cl.viz$cho$data
df.colors <- ev.df.cl.viz$cho$color
p <- EnhancedVolcano(df.viz,
    lab = rownames(df.viz),
    colCustom = df.colors,
    selectLab = sr.process.ev.cl$cho,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(df.viz$logFC),max(df.viz$logFC)),
    ylim=c(0,-log10(min(df.viz$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle(paste('GSVA -',cl.comparisons$cho,'(Protein)')) + theme(legend.position="none",plot.title = element_text(hjust = 0.5)) #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'right')

## uncomment above lines to create a manual legend for fianl figures

print(p)
# ggsave("cho_gsva_prot.svg", plot = p, width = 10, height = 8)
```

volcano plot for RNA- murine intraspecies
```{r fig.height=4, fig.width=5}
df.viz <- ev.df.cl.viz$mm$data
df.colors <- ev.df.cl.viz$mm$color
p <- EnhancedVolcano(df.viz,
    lab = rownames(df.viz),
    colCustom = df.colors,
    selectLab = sr.process.ev.cl$mu,
    boxedLabels = TRUE,
    drawConnectors = T, widthConnectors = 1.0,
    # lengthConnectors=unit(0.05, "npc"),
    # min.segment.length=0,
    arrowheads = F,
    FCcutoff = 0,
    # min.segment.length=1,
    colConnectors = 'black',
    xlab='GSVA score enrichment difference',
    ylab='-log10 (adj.p-value)',
    x = 'logFC',
    y = 'adj.P.Val',
    xlim=c(min(df.viz$logFC),max(df.viz$logFC)),
    ylim=c(0,-log10(min(df.viz$adj.P.Val))),
    pCutoff=0.05,
    shadeAlpha = 1,
    pointSize=8,subtitle='') + ggtitle(paste('GSVA -',cl.comparisons$mm,'(Protein)')) + theme(legend.position="none",plot.title = element_text(hjust = 0.5)) #+ scale_color_manual(values=sec.meta.colors.pg) + guides(color = guide_legend(title='System')) + theme(legend.position = 'right')

## uncomment above lines to create a manual legend for fianl figures

print(p)
# ggsave("mm_gsva_prot.svg", plot = p, width = 10, height = 8)
```

dotplot version of significant differential GSVA scores
```{r}
dotplot.limma <- lapply(gsva.limma.sig,function(x) {
  # remove Feizi genesets
  x.plot <- x[!grepl('f_',rownames(x)),]
  
  # x.plot <- x.plot %>% arrange(logFC)
  x.plot$log10padj <- -log10(x.plot$adj.P.Val)
  x.plot <- x.plot %>% arrange(log10padj)
  
  x.plot$term <- rownames(x.plot)

  level.order <- arrange(x.plot, logFC)
  # level.order <- arrange(x.plot, log10padj)
  # level.order$term <- rownames(level.order)
  
  x.plot <- mutate(x.plot,process=factor(term, levels=unique(level.order$term)))
  
  return(x.plot)})
#   if (is.null(x$sigRes)) {x.plot <- x %>% arrange(log2FC)} else {x.plot <- x$sigRes %>% arrange(log2FC)}
#   
#   level.order <- arrange(x.plot, log2FC)
#   x.plot <- mutate(x.plot,process=factor(process, levels=unique(level.order$process)))
# })
```


```{r fig.height=3, fig.width=4}  
for (i in names(dotplot.limma)) {
  p1 <- ggplot(dotplot.limma[[i]], aes(logFC, process)) +
    geom_point(aes(size=log10padj,color=process), alpha=0.7,show.legend = F) +
    scale_color_manual(values=sec.colmap) +
    theme(text=element_text(size=8), axis.title.y=element_blank()) +
    ggtitle(paste(i,'GSVA adj.pval<0.05')) +
    geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1) #+ scale_colour_gradient(low = "indianred3", high = "orchid4")
  print(p1)}

# scale_color_manual(values=col)
```

# identify engineering targets in enriched differential processes
```{r}
# what are the FC expr for O-glycan mucin genes
data.rna = readxl::read_excel('../data/secrecon_1-s2.0-S1096717624000521-mmc3.xlsx',sheet='Transcriptome DE')
data.prot = readxl::read_excel('./data/secrecon_1-s2.0-S1096717624000521-mmc3.xlsx',sheet='Proteome DE proteins')
```

```{r}
# get named and ranked vector of l2FC
fc.rna <- data.rna$`Fold-Change CHO/Murine`
names(fc.rna) <- toupper(data.rna$Mmus_Genes)

fc.prot <- log2(data.prot$`FC PCD/CHO`)
names(fc.prot) <- toupper(data.prot$Mmus_Genes)

## convert CHO/murine FC to l2FC and invert to murine/CHO

## convert FC to l2FC
### -FC was calculated as -1/FC according to methods --> convert it back to just FC
l2fc.rna <- sapply(fc.rna,function(x) {
  # adjust negative FC to cho/murine FC
  if (x < 0) {x.adj <- -1/x} else {x.adj <- x}
  
  # invert FC to murine/CHO FC
  # x.fc <- x.adj^-1 
  return(log2(x.adj))
})
l2fc.rna <- l2fc.rna[order(l2fc.rna)]

l2fc.prot <- fc.prot[order(fc.prot)]

# head(fc.rna)
head(l2fc.rna)
head(l2fc.prot)
```

```{r}
l2fc.rna.sec <- l2fc.rna[intersect(names(l2fc.rna),unlist(sec.db))]
l2fc.pro.sec <- l2fc.prot[intersect(names(l2fc.prot),unlist(sec.db))]

length(l2fc.rna.sec)
length(l2fc.pro.sec)

l2fc.rp.sec.g <- intersect(names(l2fc.rna.sec),names(l2fc.pro.sec))

l2fc.rp.sec <- data.frame(cbind(l2fc.rna.sec[l2fc.rp.sec.g],l2fc.pro.sec[l2fc.rp.sec.g]))
colnames(l2fc.rp.sec) <- c('rna.log2FC','pro.log2FC')
ggscatter(l2fc.rp.sec,x='rna.log2FC',y='pro.log2FC') + stat_cor(method='spearman',label.y.npc='top',size=3) + font('xy.text',size=8) + ggtitle(paste(length(l2fc.rp.sec.g),'secRecon genes'))
```

```{r}
# for the secrecon genes that are DE in both RNA/PROT, highlight the consistent upregulated terms and label the ones that are similarly differentially expressed

sr.process.ev <- c('ERpQC', 'Vesicle budding', 'PERK pathway', 'Co-translational translocation', 'N-glycan','Hydroxylysine','other')

# assign unique colorpalette to metadata
sec.ev.meta.colors.pg <- c('red','green','cornflowerblue','purple','darkgreen','darkkhaki','grey') #c('blue3','orange','green','darkkhaki','skyblue','yellow','red')

names(sec.ev.meta.colors.pg) = sr.process.ev 
sec.ev.meta.colors.pg
```

```{r}
# generate manual legend

# Sample data
sdata <- data.frame(
  x = rnorm(14),
  y = rnorm(14),
  group = factor(rep(1:7, each=2), labels = sr.process.ev))

# Create scatter plot with custom legend
ggplot(sdata, aes(x = x, y = y, color = group)) +
  geom_point(size = 3) +
  scale_color_manual(
    values = sec.ev.meta.colors.pg,
    name = "Process"  # Custom legend title
  ) +
  theme_minimal()
```


```{r}
# label consistent genes, engineering targets in scatter plot
sr.process.ev.genes <- unlist(sec.db[sr.process.ev])

sr.process.ev.genes.color <- c()
# sr.process.ev.genes.color.process <- c()
for (g in l2fc.rp.sec.g) {
  if (g %in% sr.process.ev.genes) {
    # return matched process
    g.process <- names(Filter(function(x) g %in% x, sec.db[sr.process.ev]))
    # print(length(g.process))
    sr.process.ev.genes.color = append(sr.process.ev.genes.color,sec.ev.meta.colors.pg[[g.process[[1]]]])

  } else {sr.process.ev.genes.color = append(sr.process.ev.genes.color,'grey')

}}

names(sr.process.ev.genes.color) <- l2fc.rp.sec.g


l2fc.rp.sec.g.targets <- c()

for (g in sr.process.ev.genes) {
  if (g %in% l2fc.rp.sec.g) {
    # print(g)
    if (sign(l2fc.rna.sec[[g]]) == sign(l2fc.pro.sec[[g]])) {
      l2fc.rp.sec.g.targets = append(l2fc.rp.sec.g.targets,g)
  }}}
l2fc.rp.sec.g.targets
```


```{r fig.height=3, fig.width=4}
# Create scatter plot
l2fc.rp.sec.ev <- l2fc.rp.sec
l2fc.rp.sec.ev$gene <- rownames(l2fc.rp.sec.ev)
l2fc.rp.sec.ev$process <- sr.process.ev.genes.color

ggplot(l2fc.rp.sec.ev, aes(x = rna.log2FC, y = pro.log2FC)) +
  geom_point(aes(color = factor(names(sr.process.ev.genes.color))),
                 size = 4,alpha=0.8) +  # Color points by label
  scale_color_manual(values = sr.process.ev.genes.color) +  # Set custom colors  # Add labels for specific points
  theme_minimal() + theme(legend.position="none",
                          axis.text.x =element_text(size=12),
                          axis.text.y =element_text(size=12),
                          axis.title.x =element_text(size=12),
                          axis.title.y =element_text(size=12)) + 
  geom_segment(aes(x=0.58,xend=max(l2fc.rp.sec$rna.log2FC)+1,y=0.58,yend=0.58),color='darkgrey',linetype='longdash') +
  geom_segment(aes(x=0.58,xend=0.58,y=0.58,yend=max(l2fc.rp.sec$pro.log2FC)+1),color='darkgrey',linetype='longdash') +
  geom_segment(aes(x=-0.58,xend=min(l2fc.rp.sec$rna.log2FC)-1,y=-0.58,yend=-0.58),color='darkgrey',linetype='longdash') +
  geom_segment(aes(x=-0.58,xend=-0.58,y=-0.58,yend=min(l2fc.rp.sec$pro.log2FC)-1),color='darkgrey',linetype='longdash') +
  geom_hline(yintercept =0,color='black') +
  geom_vline(xintercept =0,color='black') +
  geom_label_repel(
        data = l2fc.rp.sec.ev[l2fc.rp.sec.g.targets,],
        aes(label = l2fc.rp.sec.ev[l2fc.rp.sec.g.targets,'gene']),
        box.padding = 0.3,     # Adjust padding around labels
        point.padding = 0.2,
        # size = labSize,
        segment.color = 'grey10',
        segment.size = 0.5,
        # arrow = arr,
        colour = 'black',
        fontface = 'plain',
        parse = F,
        na.rm = TRUE,
        direction = 'both',
        max.overlaps = 15,
        min.segment.length = 0) +
  xlab('RNA log2FC') + ylab('PROTEIN log2FC') 
```


```{r}
# save.image("../data/2_gsva_limma.RData")
```

